---
title: 'exercise 1: airlines data analysis'
date: "2023-09-06"
output:
  html_notebook: default
  word_document: default
  pdf_document: default
---

```{r}
library(nycflights13)
library(tidyverse)
library(janitor)
library(here)
```

# When executing this code we are defining flights as the variable df1. View creates a new window and displays the data that way. Glimpse prints the flights data in the command window. Print prints the data in a table below the code inside the notebook.

```{r}
df1 = flights
View(flights)
glimpse(flights)
print(flights, width = Inf)
```

```{r}
df1 |>
  filter(dest == "IAH") |> 
  group_by(year, month, day) |> 
  summarize(
    arr_delay = mean(arr_delay, na.rm = TRUE)
  )
```

```{r}
df2 = df1|>
  filter(month == 9, origin == "MIA")
```

```{r}
df3 = df1|>
  filter(month == 9, dest == "MIA") 
```

```{r}
df4 = df1|>
  filter(month == 1, dest == "MIA")
```

```{r}
df5 = df1 |> 
  filter(month %in% c(7, 8, 9), dest %in% c("ORD", "MDW"))
```

```{r}
df5a=df1 |>
   filter(dest %in% c("ORD", "MDW")) |>
   filter(time_hour>="2013-06-21", time_hour<="2013-09-21")
```

```{r}
df3b = df3 |> distinct(flight)
```

```{r}
df3c = df3b |> arrange(desc(flight))
```

```{r}
df6 = df1 |> filter(dest == "MIA", flight == 83)
```

```{r}
df1 |> filter(flight == 83) |> ggplot(aes(x = dep_delay, y = arr_delay)) +
  geom_point() + 
  geom_smooth()
```

# Start of work done after mob-programming (work due Mon):

## Splitting up flight 83 data into quarters (3-month chunks) to visualize whether different times of year result in different amounts of delay.

```{r}
df1 |> filter(flight == 83) |> filter(month %in% c(1,2,3)) |> ggplot(aes(x = dep_delay, y = arr_delay)) +
  geom_point() + 
  geom_smooth()
```

```{r}
df1 |> filter(flight == 83) |> filter(month %in% c(4,5,6)) |> ggplot(aes(x = dep_delay, y = arr_delay)) +
  geom_point() + 
  geom_smooth()
```

```{r}
df1 |> filter(flight == 83) |> filter(month %in% c(7,8,9)) |> ggplot(aes(x = dep_delay, y = arr_delay)) +
  geom_point() + 
  geom_smooth()
```

```{r}
df1 |> filter(flight == 83) |> filter(month %in% c(10,11,12)) |> ggplot(aes(x = dep_delay, y = arr_delay)) +
  geom_point() + 
  geom_smooth()
```

# Start of work for 9/11 asynchronous class

## The visualization of origin points for Miami flights
```{r}
df1 |>
  filter(dest=="MIA") |>
  count(origin, sort = TRUE)
```

## New visualization of flight deltas


```{r}
df7 = df1 |>
  filter(dest == "MIA", origin == "LGA") |>
  mutate(flt_delta = arr_delay - dep_delay)
```

```{r}
df7 |> ggplot(aes(x=dep_delay, y=flt_delta)) + geom_point(alpha = .1)
```

```{r}
df7 |> 
  ggplot(aes(x=dep_delay, y=flt_delta)) + geom_point(alpha = .1) +
  geom_hline(aes(yintercept = mean(flt_delta, na.rm=TRUE)))
```

### Is departure delay affected by the time of year?
```{r}
df7 |> ggplot(aes(x=time_hour, y=dep_delay)) + geom_point(alpha = .1)
```


```{r}
df7 |> ggplot(aes(x = time_hour, y=dep_delay)) + geom_point(alpha = .1) +
  stat_smooth() + ylim(-25, 120)
```

### Does the departure delay change across time of day?
```{r}
df7 |>
  ggplot(aes(x = hour + minute/60, y = dep_delay)) +
  geom_point(alpha = .1) +
  stat_smooth() +
  ylim(-25, 120)
```


```{r}
df7 |>
  mutate(day_of_week = weekdays(time_hour)) |>
  ggplot(aes(x = hour + minute/60, y = dep_delay, color = day_of_week)) +
  geom_point(alpha = .1) +
  stat_smooth() +
  ylim(-25, 120)
```


```{r}
df7 |>
  mutate(day_of_week = weekdays(time_hour)) |>
  ggplot(aes(x = hour + minute/60, y = dep_delay, color = day_of_week)) +
  geom_point(alpha = .1) +
  stat_smooth() +
  ylim(-20, 40) +
  facet_wrap(~day_of_week)
```
