---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

### Loading packages
```{r}
library(tidyverse)
library(here)
library(janitor)
```

### Reading in data files
```{r}
thisfile_1 = here("data_raw", "DL_SelectFields_2023.zip")
df2023 = read_csv(thisfile_1) %>% clean_names()

thisfile_2 = here("data_raw", "DL_SelectFields_2022.zip")
df2022 = read_csv(thisfile_2) %>% clean_names

thisfile_3 = here("data_raw", "DL_SelectFields_2021.zip")
df2021 = read_csv(thisfile_3) %>% clean_names

thisfile_4 = here("data_raw", "DL_SelectFields_2020.zip")
df2020 = read_csv(thisfile_4) %>% clean_names

thisfile_5 = here("data_raw", "DL_SelectFields_2019.zip")
df2019 = read_csv(thisfile_5) %>% clean_names

thisfile_6 = here("data_raw", "DL_SelectFields_2018.zip")
df2018 = read_csv(thisfile_6) %>% clean_names

thisfile_7 = here("data_raw", "DL_SelectFields_2017.zip")
df2017 = read_csv(thisfile_7) %>% clean_names
```

### merging data files into one dataframe
```{r}
dfc=df2017 |>
bind_rows(df2018)  |>
bind_rows(df2019)  |>
bind_rows(df2020)  |>
bind_rows(df2021)  |>
bind_rows(df2022)  |>
bind_rows(df2023)
```

### filtering to just LaGuardia flights
```{r}
dfc1=dfc  |>
  filter(passengers > 0, origin=="LGA" | dest == "LGA")
```

### plotting by month
```{r}
dfc1 |> mutate(month_year = year + (month/12)) |> 
  ggplot(aes(month_year)) + geom_bar(aes(weight = passengers)) 
```
### filtering to only LaGuardia flights with an origin or destination of Miami
```{r}
dfc2 = dfc1 |>  
  filter(dest=="MIA" | origin=="MIA")
  
```

### Plotting by month and origin
```{r}
dfc2 |> 
  mutate(month_year = year + (month/12)) |>
  ggplot(aes(month_year, fill=origin)) + geom_bar(aes(weight = passengers, color=origin), position = 'dodge') 
```
# Start of post-class work

### Grouping by origin and month, summarizing by a sum of passengers to get one row per month
```{r}
dfc3 = dfc2 |> 
  group_by(origin, month) |>
  summarize(passenger_month = sum(passengers))
```
### Adding a total passengers variable to get a cumulative sum
```{r}
dfc4 = dfc3 |> mutate(total_passengers = cumsum(passenger_month))
```

```{r}
dfc5 = pivot_wider(dfc4, names_from = origin, values_from = c("passenger_month", "total_passengers"))
```

### Pivoting table wider
```{r}
dfc6 = dfc5 |>
  mutate(month_diff = passenger_month_MIA - passenger_month_LGA, cum_diff = total_passengers_MIA - total_passengers_LGA)
```

### Plotting difference between Miami and LaGuardia flights by month
```{r}
dfc6 |> ggplot(aes(month)) + geom_bar(aes(weight = month_diff))
```

## Now plotting between different airports: Houston and Los Angeles
### filtering to just LAX flights
```{r}
dfc1a=dfc  |>
  filter(passengers > 0, origin=="LAX" | dest == "LAX")
```

### plotting by month
```{r}
dfc1a |> mutate(month_year = year + (month/12)) |> 
  ggplot(aes(month_year)) + geom_bar(aes(weight = passengers)) 
```

### filtering to only LAX flights with an origin or destination of Jackson
```{r}
dfc2a = dfc1a |>  
  filter(dest=="MEM" | origin=="MEM")
  
```

### Plotting by month and origin
```{r}
dfc2a |> 
  mutate(month_year = year + (month/12)) |>
  ggplot(aes(month_year, fill=origin)) + geom_bar(aes(weight = passengers, color=origin), position = 'dodge') 
```

### Grouping by origin and month, summarizing by a sum of passengers to get one row per month
```{r}
dfc3a = dfc2a |> 
  group_by(origin, month) |>
  summarize(passenger_month = sum(passengers))
```
### Adding a total passengers variable to get a cumulative sum
```{r}
dfc4a = dfc3a |> mutate(total_passengers = cumsum(passenger_month))
```

```{r}
dfc5a = pivot_wider(dfc4a, names_from = origin, values_from = c("passenger_month", "total_passengers"))
```

### Pivoting table wider
```{r}
dfc6a = dfc5a |>
  mutate(month_diff = passenger_month_LAX - passenger_month_MEM, cum_diff = total_passengers_LAX - total_passengers_MEM)
```

### Plotting difference between Dallas and LAX flights by month
```{r}
dfc6a |> ggplot(aes(month)) + geom_bar(aes(weight = month_diff))
```


